# Model


## Aim 1: Exercise

**Rest versus exhaustive exercise**

Estimate model.
Specify the correlation structure using `cs`.
Use `corSymm`, *general correlation matrix, with no additional structure*.

```{r lmeDiagnosticAim1}
fixed <- formula(logValue ~
                   genotype +
                   activity +
                   metabolite +
                   genotype * activity +
                   genotype * metabolite +
                   activity * metabolite +
                   activity * metabolite * genotype)
random <- formula(~ 1 | id)
ctrl <- lmeControl(opt = "optim",
                   maxIter = 500, msMaxIter = 500)
cs <-
  corSymm(form = random, fixed = FALSE) %>%
  Initialize(data = D1)
Dim(cs)
M1 <- D1 %>% lme(fixed, data = ., random = random, correlation = cs, control = ctrl)
M1 %>% plot
M1 %>% ranef %>% plot
M1 %>% anova %>% kable
M1 %>% tidy(effects = "fixed") %>% kable
M1 %>% tidy(effects = "fixed") %>% write.csv(file = "../data/processed/lmeFixedCoefAim1.csv", row.names = FALSE)
```

Calculate contrasts of *Exercise vs Rest* given metabolite and genotype.

```{r}
Ftests <- rbind(contrastGenotype("WT", D1, "activity", "Exercise"),
                contrastGenotype("KO", D1, "activity", "Exercise"))
Ftests %>% kable
Ftests %>% write.csv(file = "../data/processed/contrastsAim1.csv", row.names = FALSE)
```

Click [link to figure](../figures/plotDataAim1.png).


## Aim 2: Chow

**MCT chow versus experimental chow with triheptanoin**

Estimate model.
Specify the correlation structure using `cs`.
Use `corSymm`, *general correlation matrix, with no additional structure*.

```{r lmeDiagnosticAim2}
rm(M)
fixed <- formula(logValue ~
                   genotype +
                   chow +
                   metabolite +
                   genotype * chow +
                   genotype * metabolite +
                   chow * metabolite +
                   chow * metabolite * genotype)
cs <-
  corSymm(form = random, fixed = FALSE) %>%
  Initialize(data = D2)
Dim(cs)
M2 <- D2 %>% lme(fixed, data = ., random = random, correlation = cs, control = ctrl)
M2 %>% plot   
M2 %>% ranef %>% plot
M2 %>% anova %>% kable
M2 %>% tidy(effects = "fixed") %>% kable
M2 %>% tidy(effects = "fixed") %>% write.csv(file = "../data/processed/lmeFixedCoefAim2.csv", row.names = FALSE)
```

Calculate contrasts of *Yellow (C8) vs White (C7)* given metabolite and genotype.

```{r}
Ftests <- rbind(contrastGenotype("WT", D2, "chow", "Yellow (C8)"),
                contrastGenotype("KO", D2, "chow", "Yellow (C8)"))
Ftests %>% kable
Ftests %>% write.csv(file = "../data/processed/contrastsAim2.csv", row.names = FALSE)
```

Click [link to figure](../figures/plotDataAim2.png).


Save `lme` objects for interactive work.

```{r}
save(M1, M2, file = "../data/processed/lmeObjects.RData")
```
