# Model


## Methods

A mixed linear effects model was estimated for each aim.
Metabolomic values were log-transformed, then normalized.
Fixed effects for Aim 1 were activity, genotype, and metabolite.
Fixed effects for Aim 2 were chow, genotype, and metabolite.
All 2-way and 3-way interactions between fixed effects were included in the models.
Animal ID was the random effect.
A general correlation structure was assumed.
Estimates for the contrasts comparing exercise versus rest for Aim 1, and yellow C8 chow versus white C7 chow for Aim 2, for each combination of genotype and metabolite are presented.
P-values were adjusted to control the false discovery rate, the expected proportion of false discoveries amongst the rejected hypotheses.
The data was analyzed using `r R.Version()$version.string` and the `nlme` package version `r packageVersion("nlme")`.


## References

Benjamini, Y., and Hochberg, Y.
(1995).
Controlling the false discovery rate: a practical and powerful approach to multiple testing.
*Journal of the Royal Statistical Society Series B* 57, 289–300.

```{r}
citation()
citation("nlme")
```


## Aim 1: Exercise

**Rest versus exhaustive exercise**

Estimate model.
Specify the correlation structure using `cs`.
Use `corSymm`, *general correlation matrix, with no additional structure*.

```{r lmeDiagnosticAim1}
fixed <- formula(zLogValue ~
                   genotype +
                   activity +
                   metabolite +
                   genotype * activity +
                   genotype * metabolite +
                   activity * metabolite +
                   activity * metabolite * genotype)
random <- formula(~ 1 | id)
ctrl <- lmeControl(opt = "optim",
                   maxIter = 500, msMaxIter = 500)
cs <-
  corSymm(form = random, fixed = FALSE) %>%
  Initialize(data = D1)
Dim(cs)
M1 <- D1 %>% lme(fixed, data = ., random = random, correlation = cs, control = ctrl)
M1 %>% plot
M1 %>% ranef %>% plot
M1 %>% anova %>% kable
M1 %>% tidy(effects = "fixed") %>% kable
M1 %>% tidy(effects = "fixed") %>% write.csv(file = "../data/processed/lmeFixedCoefAim1.csv", row.names = FALSE)
```

Calculate contrasts of *Exercise vs Rest* given metabolite and genotype.

```{r}
metabolites <- c("3-HYDROXYBUTYRIC",
                 "arginine",
                 "CITRIC",  # Returns "Singularity in backsolve at level 0, block 1" error
                 "FUMARIC",
                 "glutamine",
                 "isoleucine",
                 "LACTIC",
                 "LCAC total",  # Returns "Singularity in backsolve at level 0, block 1" error
                 "leucine",
                 "MALIC",
                 "MCAC Total",
                 "METHYLSUCCINIC",
                 "PYRUVIC_P2P",
                 "SUCCINIC-2",
                 "valine")
Ftests <- runClusters(D1, metabolites, fixed, "activity", "Exercise", ctrl)
Ftests %>% kable
Ftests %>% write.csv(file = "../data/processed/contrastsAim1.csv", row.names = FALSE)
```

Click [link to figure](../figures/plotDataAim1.png).


## Aim 2: Chow

**MCT chow versus experimental chow with triheptanoin**

Estimate model.
Specify the correlation structure using `cs`.
Use `corSymm`, *general correlation matrix, with no additional structure*.

```{r lmeDiagnosticAim2}
rm(M)
fixed <- formula(zLogValue ~
                   genotype +
                   chow +
                   metabolite +
                   genotype * chow +
                   genotype * metabolite +
                   chow * metabolite +
                   chow * metabolite * genotype)
cs <-
  corSymm(form = random, fixed = FALSE) %>%
  Initialize(data = D2)
Dim(cs)
M2 <- D2 %>% lme(fixed, data = ., random = random, correlation = cs, control = ctrl)
M2 %>% plot   
M2 %>% ranef %>% plot
M2 %>% anova %>% kable
M2 %>% tidy(effects = "fixed") %>% kable
M2 %>% tidy(effects = "fixed") %>% write.csv(file = "../data/processed/lmeFixedCoefAim2.csv", row.names = FALSE)
```

Calculate contrasts of *Yellow (C8) vs White (C7)* given metabolite and genotype.

```{r}
metabolites <- c("3-HYDROXYBUTYRIC",
                 "arginine",
                 "CITRIC",
                 "FUMARIC",
                 "glutamine",
                 "isoleucine",
                 "LACTIC",
                 "LC even AC total",
                 "LC odd AC total",
                 "leucine",
                 "MALIC",
                 "MCAC total",
                 "METHYLSUCCINIC",
                 "SUCCINIC-2",
                 "valine")
ctrl <- lmeControl(opt = "optim",
                   maxIter = 5000, msMaxIter = 5000,
                   tolerance = 1e-6, niterEM = 25, msMaxEval = 200, msTol = 1e-4)
Ftests <- runClusters(D2, metabolites, fixed, "chow", "Yellow (C8)", ctrl)
Ftests %>% kable
Ftests %>% write.csv(file = "../data/processed/contrastsAim2.csv", row.names = FALSE)
```

Click [link to figure](../figures/plotDataAim2.png).


## Save

Save `lme` objects for interactive work.

```{r}
save(M1, M2, file = "../data/processed/lmeObjects.RData")
```
