# Model


## Aim 1: Exercise

**Rest versus exhaustive exercise**

Estimate model.
Specify the correlation structure using `cs`.
Use `corSymm`, *general correlation matrix, with no additional structure*.

```{r lmeDiagnosticAim1}
fixed <- formula(logValue ~
                   genotype +
                   activity +
                   metabolite +
                   genotype * activity +
                   genotype * metabolite +
                   activity * metabolite +
                   activity * metabolite * genotype)
random <- formula(~ 1 | id)
ctrl <- lmeControl(opt = "optim",
                   maxIter = 500, msMaxIter = 500)
cs <-
  corSymm(form = random, fixed = FALSE) %>%
  Initialize(data = D1)
Dim(cs)
M <- lme(fixed, data = D1, random = random, correlation = cs, control = ctrl)
plot(M)
anova(M)
summary(M)
```


## Aim 2: Chow

**MCT chow versus experimental chow with triheptanoin**

Estimate model.
Specify the correlation structure using `cs`.
Use `corSymm`, *general correlation matrix, with no additional structure*.

```{r lmeDiagnosticAim2}
rm(M)
fixed <- formula(logValue ~
                   genotype +
                   chow +
                   metabolite +
                   genotype * chow +
                   genotype * metabolite +
                   chow * metabolite +
                   chow * metabolite * genotype)
cs <-
  corSymm(form = random, fixed = FALSE) %>%
  Initialize(data = D2)
Dim(cs)
M <- lme(fixed, data = D2, random = random, correlation = cs, control = ctrl)
plot(M)
anova(M)
summary(M)
```
