# Simulation

Load packages.

```{r}
library(data.table)
library(magrittr)
library(nlme)
library(ggplot2)
```

Source functions.

```{r}
source("../lib/R/library.R")
```


## Build simulation data

Set simulation parameters.

* `genotype` is wild type `WT` or knockout `KO`
* `activity` is `rest` or `exhaustive`
* `n` is the number of mice per (`genotype`, `activity`)
* `m` is the number of metabolites
* `j` is number of metabolite types
* `k` is number of metabolites per type
* `rho1` is the within-metabolite correlation
* `rho2` is the between-metabolite correlation

```{r}
genotype <- factor(c("KO", "WT"), levels = c("WT", "KO"))
activity <- factor(c("Exercise", "Rest"), levels = c("Rest", "Exercise"))
n <- 10
m <- 300
j <- 15
k <- m / j
rho1 <- 0.5
rho2 <- 0.2
```

Build correlated metabolite measurements.
Impose a symmetric correlation structure on them.
Assume metabolites are correlated within metabolite type with $\rho_1 = $ `r rho1`.

```{r}
id <- factor(seq(1, n * length(genotype) * length(activity)))
R <- diag(k)
R[R == 0] <- rho1
X <- list()
for (i in 1:j) {
  X[[i]] <- simulateCorrData(R, n, prefix = LETTERS[i])$data.table
}
D <-
  data.table(id, genotype = rep(genotype, each = n)) %>%
  setorder(genotype, id) %>%
  .[, activity := rep(activity, each = n)] %>%
  setorder(id) %>%
  cbind(X[[1]], X[[2]], X[[3]], X[[4]], X[[5]],
        X[[6]], X[[7]], X[[8]], X[[9]], X[[10]],
        X[[11]], X[[12]], X[[13]], X[[14]], X[[15]]) %>%
  melt(id.vars = c("id", "genotype", "activity"),
       variable.name = "metabolite",
       value.name = "y") %>%
  .[, metaboliteType := factor(substr(metabolite, 1, 1))] %>%
  .[genotype == "KO" & metaboliteType == "B", y := y + 1.0 + rnorm(length(id))] %>%
  .[activity == "Exercise" & metaboliteType == "C", y := y - 1.0 + rnorm(length(id))] %>%
  setorder(id, genotype, activity, metaboliteType, metabolite)
str(D)
```

Build the `r print(m)` $\times$ `r print(m)` correlation structure.
Assume metabolties of different types are correlated with $\rho_2 = $ `r rho2`.

```{r}
R
bigR <- matrix(rep(rho2, m * m), nrow = m)
dim(bigR)
for (i in 1:j) {
  x <- (i - 1) * k + 1:k
  bigR[x, x] <- R
}
random <- formula(~ 1 | id)
cs <-
  corSymm(value = pdSymm(bigR)[lower.tri(bigR)], form = random, fixed = TRUE) %>%
  Initialize(data = D)
Dim(cs)
# summary(cs)
```

Define the fixed effects part of the model.

```{r}
fixed <- formula(y ~ genotype +
                       activity +
                       metaboliteType +
                       genotype * metaboliteType +
                       activity * metaboliteType)
```

Estimate model.
Default to `correlation = NULL`, corresponding to no within-group correlations.

```{r}
M <- lme(fixed, data = D, random = random, correlation = NULL)
summary(M)
```

Estimate model.
Specify the correlation structure.

```{r}
M <- lme(fixed, data = D, random = random, correlation = cs)
summary(M)
```
