# Subanalysis

Test genotype, KO vs WT, within metabolite.

We want to be able to make statements such as

> metabolite X was significantly lower among the WT mice compared to the KO
> mice fed C7 but no difference was detected in WT and KO mice fed C8

using the subanalysis models.


## Aim 1

Click [link to figure](../figures/plotDataAim1Subanalysis.svg).

```{r plotDataAim1Subanalysis, results = "asis", fig.height = 6, fig.width = 10}
kable(summarizeOutcome(D1))
D1 %>%
  ggplot +
  aes(x = zLogValue, y = metabolite, color = genotype, fill = genotype) +
  geom_jitter(alpha = 1/2) +
  facet_wrap(~ activity) +
  scale_x_continuous("Normalized log-transformed values") +
  scale_y_discrete("Metabolite") +
  scale_color_brewer("Genotype", palette = "Set2") +
  scale_fill_brewer("Genotype", palette = "Set2") +
  theme_bw()
ggsave("../figures/plotDataAim1Subanalysis.svg")
```

Use same model as main analysis in [model.Rmd](model.Rmd).

```{r}
fixed <- formula(zLogValue ~
                   genotype +
                   activity +
                   metabolite +
                   genotype * activity +
                   genotype * metabolite +
                   activity * metabolite +
                   activity * metabolite * genotype)
random <- formula(~ 1 | id)
ctrl <- lmeControl(opt = "optim",
                   maxIter = 500, msMaxIter = 500,
                   tolerance = 1e-6, niterEM = 25, msMaxEval = 200, msTol = 1e-7)
cs <-
  corSymm(form = random, fixed = FALSE) %>%
  Initialize(data = D1)
runClusters <- function (df, metabolites, fixed, xvar, contrastValue, ctrl) {
  require(magrittr)
  require(dplyr)
  require(doParallel)
  require(data.table)
  activities <- levels(df$activity)
  lookup <-
    expand.grid(metabolites, activities, stringsAsFactors = FALSE) %>%
    data.frame %>%
    rename(metabolite = Var1, activity = Var2)
  n <- nrow(lookup)
  cl <- makeCluster(15)
  registerDoParallel(cl)
  L <- foreach (i = 1:n) %dopar% {
    require(magrittr)
    require(dplyr)
    require(nlme)
    dfi <-
      df %>%
      mutate(metabolite = relevel(metabolite, lookup$metabolite[i])) %>%
      mutate(activity = relevel(activity, lookup$activity[i]))
    random <- formula(~ 1 | id)
    cs <- corSymm(form = random, fixed = FALSE) %>% Initialize(data = dfi)
    M <- dfi %>% lme(fixed, data = ., random = random, correlation = cs, control = ctrl)
    M %>%
      anova(Terms = xvar) %>%
      data.frame(contrast = contrastValue,
                 metabolite = lookup$metabolite[i],
                 activity = lookup$activity[i],
                 beta = M %>% fixef %>% .[names(.) == paste0(xvar, contrastValue)],
                 se = M %>% summary %>% .$tTable %>% data.frame %>%
                        select(matches("Std.Error")) %>%
                        .[row.names(.) == paste0(xvar, contrastValue), ],
                 .) %>%
      mutate(lowerCL = beta + qnorm(0.025) * se,
             upperCL = beta + qnorm(0.975) * se)
  }
  stopCluster(cl)
  rbindlist(L) %>%
    mutate(p.adjustBH = p.adjust(p.value, method = "BH"),
           sig = p.adjustBH < 0.05)
}
```

Calculate contrasts of *WT vs KO* given metabolite and activity.

```{r}
metabolites <- c("3-HYDROXYBUTYRIC",
                 "arginine",
                 "CITRIC",  # Returns "Singularity in backsolve at level 0, block 1" error
                 "FUMARIC",
                 "glutamine",
                 "isoleucine",
                 "LACTIC",
                 "LCAC total",  # Returns "Singularity in backsolve at level 0, block 1" error
                 "leucine",
                 "MALIC",
                 "MCAC Total",
                 "METHYLSUCCINIC",
                 "PYRUVIC_P2P",
                 "SUCCINIC-2",
                 "valine")
Ftests <- runClusters(D1, metabolites, fixed, "genotype", "KO", ctrl)
Ftests %>% kable
Ftests %>% write.csv(file = "../data/processed/contrastsAim1Subanalysis.csv", row.names = FALSE)
```


## Aim 2

Click [link to figure](../figures/plotDataAim2Subanalysis.svg).

```{r plotDataAim2Subanalysis, results = "asis", fig.height = 6, fig.width = 10}
kable(summarizeOutcome(D2))
D2 %>%
  ggplot +
  aes(x = zLogValue, y = metabolite, color = genotype, fill = genotype) +
  geom_jitter(alpha = 1/2) +
  facet_wrap(~ chow) +
  scale_x_continuous("Normalized log-transformed values") +
  scale_y_discrete("Metabolite") +
  scale_color_brewer("Genotype", palette = "Set2") +
  scale_fill_brewer("Genotype", palette = "Set2") +
  theme_bw()
ggsave("../figures/plotDataAim2Subanalysis.svg")
```

Use same model as main analysis in [model.Rmd](model.Rmd).

```{r}
fixed <- formula(zLogValue ~
                   genotype +
                   chow +
                   metabolite +
                   genotype * chow +
                   genotype * metabolite +
                   chow * metabolite +
                   chow * metabolite * genotype)
cs <-
  corSymm(form = random, fixed = FALSE) %>%
  Initialize(data = D2)
runClusters <- function (df, metabolites, fixed, xvar, contrastValue, ctrl) {
  require(magrittr)
  require(dplyr)
  require(doParallel)
  require(data.table)
  chows <- levels(df$chow)
  lookup <-
    expand.grid(metabolites, chows, stringsAsFactors = FALSE) %>%
    data.frame %>%
    rename(metabolite = Var1, chow = Var2)
  n <- nrow(lookup)
  cl <- makeCluster(15)
  registerDoParallel(cl)
  L <- foreach (i = 1:n) %dopar% {
    require(magrittr)
    require(dplyr)
    require(nlme)
    dfi <-
      df %>%
      mutate(metabolite = relevel(metabolite, lookup$metabolite[i])) %>%
      mutate(chow = relevel(chow, lookup$chow[i]))
    random <- formula(~ 1 | id)
    cs <- corSymm(form = random, fixed = FALSE) %>% Initialize(data = dfi)
    M <- dfi %>% lme(fixed, data = ., random = random, correlation = cs, control = ctrl)
    M %>%
      anova(Terms = xvar) %>%
      data.frame(contrast = contrastValue,
                 metabolite = lookup$metabolite[i],
                 chow = lookup$chow[i],
                 beta = M %>% fixef %>% .[names(.) == paste0(xvar, contrastValue)],
                 se = M %>% summary %>% .$tTable %>% data.frame %>%
                        select(matches("Std.Error")) %>%
                        .[row.names(.) == paste0(xvar, contrastValue), ],
                 .) %>%
      mutate(lowerCL = beta + qnorm(0.025) * se,
             upperCL = beta + qnorm(0.975) * se)
  }
  stopCluster(cl)
  rbindlist(L) %>%
    mutate(p.adjustBH = p.adjust(p.value, method = "BH"),
           sig = p.adjustBH < 0.05)
}
```

Calculate contrasts of *Yellow (C8) vs White (C7)* given metabolite and chow.

```{r}
metabolites <- c("3-HYDROXYBUTYRIC",
                 "arginine",
                 "CITRIC",
                 "FUMARIC",
                 "glutamine",
                 "isoleucine",
                 "LACTIC",
                 "LC even AC total",
                 "LC odd AC total",
                 "leucine",
                 "MALIC",
                 "MCAC total",
                 "METHYLSUCCINIC",
                 "SUCCINIC-2",
                 "valine")
Ftests <- runClusters(D2, metabolites, fixed, "genotype", "KO", ctrl)
Ftests %>% kable
Ftests %>% write.csv(file = "../data/processed/contrastsAim2Subanalysis.csv", row.names = FALSE)
```
