# Subanalysis

Test genotype, KO vs WT, within metabolite.

We want to be able to make statements such as

> metabolite X was significantly lower among the WT mice compared to the KO
> mice fed C7 but no difference was detected in WT and KO mice fed C8

using the subanalysis models.


## Aim 1

Click [link to figure](../figures/plotDataAim1Subanalysis.svg).

```{r plotDataAim1Subanalysis, results = "asis", fig.height = 6, fig.width = 10}
kable(summarizeOutcome(D1))
D1 %>%
  ggplot +
  aes(x = zLogValue, y = metabolite, color = genotype, fill = genotype) +
  geom_jitter(alpha = 1/2) +
  facet_wrap(~ activity) +
  scale_x_continuous("Normalized log-transformed values") +
  scale_y_discrete("Metabolite") +
  scale_color_brewer("Genotype", palette = "Set2") +
  scale_fill_brewer("Genotype", palette = "Set2") +
  theme_bw()
ggsave("../figures/plotDataAim1Subanalysis.svg")
```

Use same model as main analysis in [model.Rmd](model.Rmd).

```{r}
fixed <- formula(zLogValue ~
                   genotype +
                   activity +
                   metabolite +
                   genotype * activity +
                   genotype * metabolite +
                   activity * metabolite +
                   activity * metabolite * genotype)
random <- formula(~ 1 | id)
ctrl <- lmeControl(opt = "optim",
                   maxIter = 500, msMaxIter = 500,
                   tolerance = 1e-6, niterEM = 25, msMaxEval = 200, msTol = 1e-7)
cs <-
  corSymm(form = random, fixed = FALSE) %>%
  Initialize(data = D1)
```

Calculate contrasts of *WT vs KO* given metabolite and genotype.

```{r}
metabolites <- c("3-HYDROXYBUTYRIC",
                 "arginine",
                 "CITRIC",  # Returns "Singularity in backsolve at level 0, block 1" error
                 "FUMARIC",
                 "glutamine",
                 "isoleucine",
                 "LACTIC",
                 "LCAC total",  # Returns "Singularity in backsolve at level 0, block 1" error
                 "leucine",
                 "MALIC",
                 "MCAC Total",
                 "METHYLSUCCINIC",
                 "PYRUVIC_P2P",
                 "SUCCINIC-2",
                 "valine")
Ftests <- runClusters(D1, metabolites, fixed, "genotype", "KO", ctrl)
Ftests %>% kable
Ftests %>% write.csv(file = "../data/processed/contrastsAim1Subanalysis.csv", row.names = FALSE)
```


## Aim 2

Click [link to figure](../figures/plotDataAim2Subanalysis.svg).

```{r plotDataAim2Subanalysis, results = "asis", fig.height = 6, fig.width = 10}
kable(summarizeOutcome(D2))
D2 %>%
  ggplot +
  aes(x = zLogValue, y = metabolite, color = genotype, fill = genotype) +
  geom_jitter(alpha = 1/2) +
  facet_wrap(~ chow) +
  scale_x_continuous("Normalized log-transformed values") +
  scale_y_discrete("Metabolite") +
  scale_color_brewer("Genotype", palette = "Set2") +
  scale_fill_brewer("Genotype", palette = "Set2") +
  theme_bw()
ggsave("../figures/plotDataAim2Subanalysis.svg")
```

Use same model as main analysis in [model.Rmd](model.Rmd).

```{r}
fixed <- formula(zLogValue ~
                   genotype +
                   chow +
                   metabolite +
                   genotype * chow +
                   genotype * metabolite +
                   chow * metabolite +
                   chow * metabolite * genotype)
cs <-
  corSymm(form = random, fixed = FALSE) %>%
  Initialize(data = D2)
```

Calculate contrasts of *Yellow (C8) vs White (C7)* given metabolite and genotype.

```{r}
metabolites <- c("3-HYDROXYBUTYRIC",
                 "arginine",
                 "CITRIC",
                 "FUMARIC",
                 "glutamine",
                 "isoleucine",
                 "LACTIC",
                 "LC even AC total",
                 "LC odd AC total",
                 "leucine",
                 "MALIC",
                 "MCAC total",
                 "METHYLSUCCINIC",
                 "SUCCINIC-2",
                 "valine")
Ftests <- runClusters(D2, metabolites, fixed, "genotype", "KO", ctrl)
Ftests %>% kable
Ftests %>% write.csv(file = "../data/processed/contrastsAim2Subanalysis.csv", row.names = FALSE)
```
